---
title: "svar package"
author: "Simone Vazzoler"
date: "`r Sys.Date()`"
output: 
  rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Estimation of a VAR model

Use `estimateVAR`. The arguments of the function are:
  
  * `rets`: the multivariate time series (variables in columns, observations in rows);
  * `p`: the order of the VAR model to be estimated;
  * `penalty`: the penalty used in Least Squares (LS). Possible values are: `"ENET"`, `"SCAD"` or `"MCP"`;
  * `options`: list of options. Some will depend on the penalty and some are global.

Global options:

  * `parallel`: `TRUE` or `FALSE` (default). Parallel cross-validation (on the folds);
  * `ncores`: if `parallel = TRUE` then you must specify the number of cores used for the parallelization (default = `1`).
  * `nfolds`: number of folds to use in the cross validation (default `nfolds = 10`)

`penalty = "ENET` options:

  * `lambda`: `"lambda.min"` (default) or `"lambda.1se"`;
  * `alpha`: a value in (0,1) (default `alpha = 1`). `alpha = 1` is LASSO regression, `alpha = 0` is Ridge LS;
  * `type.measure`: `"mse"` (default) or `"mae"`;
  * `nlambda`: number of lambdas used for cross validation.

`penalty = "SCAD"` or `"MCP"` options:
  * `eps`: convergence tolerance

Output:

  * `A`: a list of length `p` containing the matrices estimated for the VAR(p) model;
  * `time`: elapsed time for the estimation;
  * `mse`: the mean square error of the cross validation;
  * `fit`: the estimated model.

### Examples

```{r, eval=FALSE}
results <- estimateVAR(rets)
```
will estimate VAR(1) process using LASSO regression on the dataset `rets`.

The command
```{r, eval=FALSE}
results <- estimateVAR(rets, p = 3, penalty = "ENET", 
                       options = list(parallel = TRUE, ncores = 5, alpha = 0.95, 
                                      type.measure = "mae", lambda = "lambda.1se"))
```
will estimate a VAR(3) model on the dataset `rets` using the penalty `"ENET"` with `alpha = 0.95` (between LASSO and Ridge). For the cross validation it will use `"mae"` (mean absolute error) insteadof mean square error and it will choose as model the one correspondent to the lambda which is at 1 std deviation from the minimum. Moreover it will parallelize the cross validation over 5 cores.

### Simulations

Use `simulateVAR`. The parameters for the function are:

  * `N`: the dimension of the process;
  * `nobs`: the number of observations of the process;
  * `rho`: the variance/covariance "intensity";
  * `sparsity`: the percentage of non zero elements in the matrix of the VAR;
  * `method`: `"normal"` or `"bimodal"`.

```{r, eval=FALSE}
sim <- simulateVAR(N = 100, nobs = 250, rho = 0.75, sparsity = 0.05, method = "normal")
```

